---
title: "6 Feature Selection by Day"
author: "Team Strikes"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
  code_folding: hide
  highlight: pygment
  theme: united
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r Check installed packages, echo = FALSE, warning=FALSE, message=FALSE}
# Creating a vector of packages used within
packages <- c(
  'Boruta',
  'caret',
  'dplyr',
  'leaps',
  'lubridate',
  'magrittr',
  'mctest',
  'psych',
  'ranger',
  'ROCR',
  'tidyverse',
  'varrank',
  'corrplot',
  'glmnet',
  'AppliedPredictiveModeling',
  'h2o',
  'lime',
  'DT'
)

# Checking for package installations on the system and installing if not found
if (length(setdiff(packages, rownames(installed.packages()))) > 0) {
  install.packages(setdiff(packages, rownames(installed.packages())))  
}

# Including the packages for use
for(package in packages){
  library(package, character.only = TRUE)
}

# Initiate h2o automl
h2o.init()

```

### Data Understandingand Aggregation

We establish meaningful spatial relationships amongst four datasets: bird strikes, bird count, flight details, and weather. We joined these data into one master dataset, aggregated daily to ensure variable cohesiveness as well as model output effectiveness.

1.	Bird Strikes –We obtained the full wildlife strike dataset from the Federal Aviation Administration (FAA) and consider it to be our primary data source. The FAA Wildlife Strike Database covers 92 attributes for 211,000 recorded observations of any wildlife strike dating back to 1990.Some key attributes include airfield, operator, number of engines, altitude, and incident date.

2.	Bird Count –To baseline our bird strike information, we obtained the total count of observed birds from an online global database, called eBird, that is maintained as part of a project from the Cornell Lab of Ornithology. We aggregated this information by date and joined relevant attributes, such as sighting longitude and latitude, type of species, and most importantly, the bird count, to our master dataset.

3.	Flight Details–To supplement our primary dataset, we obtained flight specifics from the Bureau of Transportation Statistics by the US Department of Transportation. Details include flight identification information, arrival and departure information, and time delays.

4.	Weather –We queried weather data from the National Oceanic and Atmospheric Administration by the US Department of Commerce in order to enhance our master dataset with strike conditions. Weather variables, such as cloud height, sky cover, temperature, precipitation, and wind speed, allow us to understand the correlation or effect of environmental conditions on bird strikes.

For modeling purposes, we are analyzing the data by aggregating the data by Day and Time. Multiple regression and classification techniques will be implemented to predict the strike counts or the probability of a strike respectively based on the business understanding. To be able to model the outcomes effectively, some basic analysis will be performed on the dataset to remove noise, selection bias and high variance. Below are few of the techniques we will follow to understand more about data. 

1) Validating the Non zero Variance among all the variables in the dataset. 

2) Correlations among all the continous variables.

3) Checking the variance inflation factors amoung the continuous variables for regression analysis.

4) Identifying the importance of variables based on the target variables. 


### Data Processing

Before we start modeling we aggregate the data by day and remove all the unused variables from the original dataset. Also, split the date column into day, month, year and week of year and bind the columns to the original dataset.

```{r Model Data by Day}

########### Data Pre-Processing  ###########
# Read dataset from RDS
model.data <- readRDS("KDEN_FINAL.RDS")

#Running the Code to Group the Data by Day. THis removes Rtime, Arrivals, Departures
model.data.day <- model.data %>%
  group_by(`AIRPORT ID`, DATE) %>%
  summarise(
    # We could either take max or first for Flight count
    FLIGHTCOUNT = max(`TOTAL`),
    STRIKECOUNT = sum(STRIKE),
    #Sum of Strikes, Will be Converted to Binary Strikes later
    STRIKE = sum(STRIKE),
    BIRDCOUNT = mean(BIRDCOUNT),
    TEMP = mean(TEMP, na.rm = TRUE),
    SLP = mean(SLP, na.rm = TRUE),
    `WIND ANGLE` = mean(`WIND ANGLE`, na.rm = TRUE),
    `WIND SPEED` = mean(`WIND SPEED`, na.rm = TRUE),
    `COVER` = mean(`COVER`, na.rm = TRUE),
    CLOUDH = mean(CLOUDH, na.rm = T),
    COVERH = mean(COVERH, na.rm = T),
    PRECIP = mean(PRECIP, na.rm = T),
    VIS = mean(VIS, na.rm = T)
  ) %>%
  mutate(STRIKE = ifelse(STRIKE > 0, 1, 0))

#Converting the new day data into a data frame
model.data.day <- as.data.frame(model.data.day)
model.data.day$STRIKE <- as.factor(model.data.day$STRIKE)

# Add Year, Month, Day and Weekday fields to  the dataset as factors
# Bind new cols to the original dataset
model.data.day <- cbind(
  model.data.day,
  data.frame(
    YEAR = as.factor(format(model.data.day$DATE, format = "%Y")),
    MONTH = as.factor(format(model.data.day$DATE, format = "%m")),
    DAY = as.factor(format(model.data.day$DATE, format = "%d")),
    WEEKOFYEAR = as.factor(format(model.data.day$DATE, format = "%V"))
  )
)

# Remove AirportId and Date columns
model.data.day <-
  subset(model.data.day, select = -c(`AIRPORT ID`, DATE))

# Rename Windangle and windspeed
model.data.day <-
  model.data.day %>% dplyr::rename(WANGLE = `WIND ANGLE`,
                                   WSPEED = `WIND SPEED`)

# Because of memory constraints considering only 2 years data
model.data.day <-
  model.data.day %>% filter(!model.data.day$YEAR %in% c(2013, 2019))

model.data.day <- model.data.day %>%
  mutate(STRIKE = ifelse(STRIKE == 0, "NO", "YES"))

rm(model.data)
```

### Check Non Zero Variance Predictors

Non-zero variance technique performs multiple checks on predictors and provides information on the variables which are constants or have very little variation across all the observations. nearZeroVar(nzv function) from the caret package removes 

1) Predictors that have one unique value across samples (zero variance predictors).

2) Predictors with few unique values relative to the number of observations. 

3) Large ratio of the frequency of the most common value to the frequency of the second most common value (near-zero variance predictors).

```{r nzv}

nzv <- nearZeroVar(model.data.day, saveMetrics = TRUE)

nzv[nzv[, "zeroVar"] > 0, ] # Check for zero variance predictors. None

nzv[nzv[, "zeroVar"] + nzv[, "nzv"] > 0, ] # Check for near-zero variance predictors
```

As the above code returned no results we are now certain that there is variablility in the data relative to all the samples and the variability is signification among all unique observations in each column. 

### Correlation Plots

We now look at how the features are correlated with each other and draw some conclusions based on the results. 


```{r Correlation Function}
#External Plot for Corr Plot 
source("http://www.sthda.com/upload/rquery_cormat.r")

# Consider only numeric columns or corelation plots.
corr.plot.fn <- function(corr.data) {
  corr.list <-
    rquery.cormat(corr.data, type = "flatten", graph = FALSE)
  cormat <- rquery.cormat(corr.data, graphType = "heatmap")
  rquery.cormat(corr.data, type = "full")
  corr.list$r
}

```



```{r Correlation Plot}

# Correlation Plot with all features
corr.data <-
  model.data.day %>% dplyr::select(
    STRIKECOUNT,
    WANGLE,
    WSPEED,
    TEMP,
    CLOUDH,
    COVER,
    COVERH,
    BIRDCOUNT,
    FLIGHTCOUNT,
    SLP,
    VIS,
    PRECIP
  )

corr.plot.fn(corr.data)

```

Based on the above plot, Visibility and Cloud height are positively and strongly correlated with Cover where as Cover height is moderately positively correlated with Cover. Cover height and Cloud height are strongly negatively correlated with each other. This phenomenon also showcases that Cover, Cover height and Cloud height variables are multicollinear and there is strong evidence to remove the multi-collinearity. Sea level pressure is also positively correlated with wind speed and temperature.

With the above understanding we may want to remove the highly correlated variables and check if there are any more strongly correlated variables which surface up.

```{r}

# Correlation Plot after removing highly correlated features CloudH, CoverH, Visibility
corr.data <-
  model.data.day %>% dplyr::select(
    STRIKECOUNT,
    WANGLE,
    WSPEED,
    TEMP,
    COVER,
    BIRDCOUNT,
    FLIGHTCOUNT,
    SLP,
    PRECIP
  )

corr.plot.fn(corr.data)

```

In the new correlation plot we obtained after removing the highly correlated feature Cover height, Cloud height and Visibility, we now see that there are no more stronger correlations between any variables. With this understanding, **CloudH**, **CoverH** and **Visibility** may be the variables to be eliminated but we may want to see if any other statistical techniques may provide a similar outcome. 

```{r Information Gain}
# model.hurdle<- pscl::hurdle(regression3, data = train.data, dist = "poisson", zero.dist = "binomial")
# validateAndPrintResultC(model.hurdle,test.data)
# 
# regression3<-STRIKECOUNT~
#   FLIGHTCOUNT+
  #STRIKE+
  #BIRDCOUNT
  #TEMP+
  #SLP+
  #WANGLE+
  #WSPEED+
  #COVER+
  #CLOUDH+
  #COVERH+
  #PRECIP+
  #VIS
  #YEAR+
  #MONTH+
  #DAY+
  #WEEKOFYEAR

#Cant run the hurdle since the variables are highly correlated with one another
  
```


### Variable Importance Analysis for Classification Models


```{r Random Forest Variable Importance}
# Use Random Forest variable importance technique for variable selection
# The below list has been tailored after multiple iterations
fit <- randomForest::randomForest(
  as.factor(STRIKE) ~ WSPEED + WANGLE + COVER + BIRDCOUNT + FLIGHTCOUNT + MONTH + TEMP + COVERH + CLOUDH + YEAR + VIS + SLP + PRECIP,
  data = model.data.day,
  mtry = 2,
  importance = TRUE,
  proximity = TRUE,
  do.trace = 100
)
print(fit)

varImp(fit)
randomForest::varImpPlot(fit, type = 2)
importanceOrder = order(-fit$importance)

varImp(fit)
```

### Variable Importance using Boruta

##### How does this algorithm work?

1) Firstly, it adds randomness to the given data set by creating shuffled copies of all features (which are called shadow features).

2) Then, it trains a random forest classifier on the extended data set and applies a feature importance measure (the default is Mean Decrease Accuracy) to evaluate the importance of each feature where higher means more important.

3) At every iteration, it checks whether a real feature has a higher importance than the best of its shadow features (i.e. whether the feature has a higher Z score than the maximum Z score of its shadow features) and constantly removes features which are deemed highly unimportant.

4) Finally, the algorithm stops either when all features gets confirmed or rejected or it reaches a specified limit of random forest runs.



```{r Baruta}

# Execute Boruta
var.boruta <-
  Boruta(
    as.factor(STRIKE) ~ WANGLE + WSPEED + TEMP + COVER + BIRDCOUNT + FLIGHTCOUNT + SLP + VIS + PRECIP + MONTH + CLOUDH + COVERH + YEAR + DAY,
    data = model.data.day,
    doTrace = 2
    )

# Plot importance based on the Z Scores
lz <- lapply(1:ncol(var.boruta$ImpHistory), function(i)
  var.boruta$ImpHistory[is.finite(var.boruta$ImpHistory[, i]), i])
names(lz) <- colnames(var.boruta$ImpHistory)
Labels <- sort(sapply(lz, median))
plot(
  var.boruta,
  side = 1,
  las = 2,
  labels = names(Labels),
  at = 1:ncol(var.boruta$ImpHistory),
  cex.axis = 0.7
)
final.boruta <- TentativeRoughFix(var.boruta)
getSelectedAttributes(final.boruta, withTentative = F)
boruta.df <- attStats(final.boruta)
print(boruta.df)

```

Blue boxplots correspond to minimal, average and maximum Z score of a shadow attribute. Red, yellow and green boxplots represent Z scores of rejected, tentative and confirmed attributes respectively.

### Variable Importance using Caret package

```{r Train Control & Importance}
# prepare training scheme
control <- trainControl(method = "cv", number = 2)

# train the model using Learning Vector Quantization (LVQ) model
model <-
  train(
    as.factor(STRIKE) ~ WSPEED + WANGLE + COVER + BIRDCOUNT + FLIGHTCOUNT + MONTH + TEMP + COVERH + CLOUDH + YEAR + VIS + SLP + PRECIP + DAY,
    data = model.data.day,
    method = "lvq",
    preProcess = "scale",
    trControl = control
  )

# estimate variable importance
importance <- varImp(model, useModel = FALSE)

# summarize importance
print(importance)

# plot importance
plot(importance)
```

```{r H20 AutoMl Variable Selection}

# Identify predictors and response
y <- "STRIKE"
x <- setdiff(names(model.data.day), c("STRIKE", "STRIKECOUNT", "COVERH", "SLP", "DAY", "WEEKOFYEAR", "CLOUDH"))

# For binary classification, response should be a factor
model.data.day[, y] <- as.factor(model.data.day[, y])

# Run AutoML for 10 models
aml <- h2o.automl(x = x, y = y,
                  training_frame = as.h2o(model.data.day),
                  max_models = 10,
                  seed = 123)

# View the AutoML Leaderboard
lb <- aml@leaderboard
print(lb, n = nrow(lb))

# Get model ids for all models in the AutoML Leaderboard
model_ids <- as.data.frame(lb$model_id)[,1]

# View variable importance for all the models (besides Stacked Ensemble)
for (model_id in model_ids) {
  print(model_id)
  m <- h2o.getModel(model_id)
  h2o.varimp(m)
  h2o.varimp_plot(m)
}


 ## Train explainer
 explainer <- lime(model.data.day, aml)

 ## Get explanations for a subset of samples
 explanation <- explain(train[1:5, ], explainer, n_features = 10)

 ## Plot global explanations
 plot_explanations(explanation)

 ## Plot local explanations
 plot_features(explanation)

```


```{r Classification Model Variable Selection}



model.cla <- setdiff(names(model.data.day), c("STRIKECOUNT", "COVERH", "SLP", "DAY", "WEEKOFYEAR", "CLOUDH"))

model.cla <- model.data.day %>% select(model.cla)

### Save the file
saveRDS(model.cla,"KDEN_Class_Data.RDS")

```



### Feature Selection for Regression Models


#### Variance Inflation Factor


```{r }

#VIF Function
VIF <-
  function(linear.model,
           no.intercept = FALSE,
           all.diagnostics = FALSE,
           plot = FALSE) {
    if (no.intercept == FALSE)
      design.matrix <- model.matrix(linear.model)[, -1]
    if (no.intercept == TRUE)
      design.matrix <- model.matrix(linear.model)
    if (plot == TRUE)
      mc.plot(design.matrix, linear.model$model[1])
    if (all.diagnostics == FALSE)
      output <-
        imcdiag(design.matrix,
                linear.model$model[1],
                method = 'VIF')$idiags[, 1]
    if (all.diagnostics == TRUE)
      output <-
        imcdiag(design.matrix, linear.model$model[1])
    output
  }
```


```{r VIF}
set.seed(123)
strikes.sample <-
  model.data.day[, c(
    "STRIKECOUNT",
    "WANGLE",
    "WSPEED",
    "TEMP",
    "COVER",
    "BIRDCOUNT",
    "FLIGHTCOUNT",
    "PRECIP", 
    "CLOUDH",
    "COVERH",
    "VIS",
    "SLP"
  )]

strikes.sample$STRIKECOUNT <- as.numeric(strikes.sample$STRIKECOUNT)

fit <- glm(STRIKECOUNT ~ ., data = strikes.sample)
VIF(fit, all.diagnostics = TRUE, plot = TRUE)

```


### Variable Importance using Boruta for Regression


```{r Baruta}

var.boruta <-
  Boruta(
    STRIKECOUNT ~ WANGLE + WSPEED + TEMP + COVER + BIRDCOUNT + FLIGHTCOUNT + SLP + VIS + PRECIP + MONTH + COVERH + CLOUDH + DAY + YEAR,
    data = model.data.day,
    doTrace = 2
    )

# plot(var.boruta, cex.axis = 0.8, las = 1)

lz <- lapply(1:ncol(var.boruta$ImpHistory), function(i)
  var.boruta$ImpHistory[is.finite(var.boruta$ImpHistory[, i]), i])
names(lz) <- colnames(var.boruta$ImpHistory)
Labels <- sort(sapply(lz, median))
plot(
  var.boruta,
  side = 1,
  las = 2,
  labels = names(Labels),
  at = 1:ncol(var.boruta$ImpHistory),
  cex.axis = 0.7
)
final.boruta <- TentativeRoughFix(var.boruta)
getSelectedAttributes(final.boruta, withTentative = F)
boruta.df <- attStats(final.boruta)
print(boruta.df)


```


### Stepwise feature selection using glmStepAIC


```{r}
# Fit a logistic regression model with stepwise AIC

# This code takes a very long time to execute if WeekOfYear included
glmStepAIC <-
  train(
    STRIKECOUNT ~ WANGLE + WSPEED + TEMP + COVER + BIRDCOUNT + FLIGHTCOUNT + SLP + VIS + PRECIP + MONTH + COVERH + CLOUDH + YEAR,
    data = model.data.day,
    method = "glmStepAIC",
    allowParallel = TRUE
  )
summary(glmStepAIC)
varImp(glmStepAIC)
plot(varImp(glmStepAIC))
```

```{r Random Forest Variable Importance}
# Use Random Forest variable importance technique for variable selection
# The below list has been tailored after multiple iterations
fit <- randomForest::randomForest(
  STRIKECOUNT ~ WSPEED + WANGLE + COVER + BIRDCOUNT + FLIGHTCOUNT + MONTH + TEMP + COVERH + CLOUDH + VIS + SLP + PRECIP + YEAR,
  data = model.data.day,
  mtry = 3,
  importance = TRUE,
  proximity = TRUE,
  do.trace = 100
)
print(fit)

varImp(fit)
randomForest::varImpPlot(fit, type = 2)
importanceOrder = order(-fit$importance)

varImp(fit)
```


```{r Regression Model Variable Selection}

model.reg <- setdiff(names(model.data.day), c("STRIKE", "DAY", "WEEKOFYEAR", "PRECIP"))

model.reg <- model.data.day %>% select(model.reg)

### Save the file
saveRDS(model.cla,"KDEN_Reg_Model.RDS")

```


