---
title: "Bird and Strike distribution"
author: "Project Team"
date: "17/02/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lubridate)
library(kableExtra)
data <- readRDS("~/WTFProject/Birdstrikes_Models/KDEN_FINAL_NEW.RDS")
```

### Distribution of birdcounts
The following plots exclude zeros and are limited to counts below 2,500, 5,000 and 10,000 for readability
```{r, echo=FALSE}
ggplot(data[data$BIRDCOUNT<2500 & data$BIRDCOUNT>0,], aes(x=BIRDCOUNT,fill=as.factor(STRIKE))) + 
  geom_histogram(binwidth = 200) + 
  scale_color_manual(values = c("#E69F11","#56B4E8"))

ggplot(data[data$BIRDCOUNT<5000 & data$BIRDCOUNT>0,], aes(x=BIRDCOUNT,fill=as.factor(STRIKE))) + 
  geom_histogram(binwidth = 200) + 
  scale_color_manual(values = c("#E69F11","#56B4E8"))

ggplot(data[data$BIRDCOUNT<10000 & data$BIRDCOUNT>0,], aes(x=BIRDCOUNT,fill=as.factor(STRIKE))) + 
  geom_histogram(binwidth = 200) + 
  scale_color_manual(values = c("#E69F11","#56B4E8"))

```


### Birdcounts vs. Chance of Strike
Average chance of strike for each bucket of bird counts (1,000 increments) - no clear relationship across the range, although it would seem that days with extremely high bird observations do have a meaningfully larger chance, although as the table below the chart demonstrates, this is the result of only two data points.
  
  
```{r, echo=FALSE, warning=FALSE}
options(scipen=999)
# Create buckets to group data
data <- data %>% mutate(BIRDBUCKET = cut(BIRDCOUNT, 
                                         breaks = seq(0,20000,1000), 
                                         labels=paste("up to:",cumsum(diff(seq(0,20000,1000)))),
                                         include.lowest = TRUE))
# Calculate avearge strike % per bucket
d <- data %>% mutate(RATIO = STRIKECOUNT/FLIGHTCOUNT) %>% group_by(BIRDBUCKET) %>% summarise(STRIKECHANCE=mean(RATIO))
ggplot(d, aes(x=BIRDBUCKET,y=STRIKECHANCE)) + geom_bar(stat="identity") + coord_flip()
```

The above data in table form:
```{r, echo=FALSE, warning=FALSE}
d<-data %>% mutate(RATIO = STRIKECOUNT/FLIGHTCOUNT) %>% 
  group_by(BIRDBUCKET) %>% summarise(N=n(), STRIKECHANCE=mean(RATIO), STRIKES=sum(STRIKECOUNT), FLIGHTS = sum(FLIGHTCOUNT))
kable(d) %>% kable_styling(bootstrap_options = c("striped","hover", "condensed"))
```


### Distribution of Strikes
Distribution of Strikes (first row=number of strikes, second row = number of data points with this many strikes)
```{r, echo=FALSE}
table(data$STRIKECOUNT)

ggplot(data, aes(STRIKECOUNT)) + geom_histogram(binwidth=1)
```

