---
title: "Models"
author: "Christian Endter"
date: "12/01/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)


library(MASS)
library(gbm)
library(tidyverse)
library(lubridate)
library(ggplot2)
library(pscl)
library(lme4)
library(glmmTMB)
library(dataPreparation)
library(gbm)
library(caret)
library(ROCR)
library(performanceEstimation)
library(rpart)
library(e1071)
library(plm) # Hausman test
# library(ResourceSelection)


```


```{r Data set preparation}
data <- readRDS("KDEN model data all flights.RDS")


### Set 1: Base data (adjust time filter as needed)
### - every line corresponds to a single flight with STRIKE = 1/0
d <- data %>% dplyr::filter(year(DATE) %in% 1992:2019) %>%
    mutate(WEEK = as.factor(isoweek(DATE)), 
           MONTH = as.factor(month(DATE)))
  
### Set 2: Data aggregated by hour
### - every line corresponds to a particular hour on a date with STRIKES = num strikes, and STRIKE = 1/0
d_day_hour <- d %>%
  group_by(`AIRPORT ID`,DATE,RTIME) %>%
  summarise(`FLTS TOTAL` = first(`TOTAL`), STRIKES = sum(STRIKE),
            BIRDCOUNT = first(BIRDCOUNT),
            TEMP = mean(TEMP,na.rm=TRUE), SLP = mean(SLP, na.rm=TRUE),
            `WIND ANGLE` = mean(`WIND ANGLE`, na.rm=T), `WIND SPEED`= mean(`WIND SPEED`,na.rm=TRUE),
            CLOUDH = mean(CLOUDH, na.rm=T), COVERH = mean(COVERH, na.rm=T),
            VIS = mean(VIS, na.rm=T),
            PRECIP = mean(PRECIP, na.rm=T)) %>%
  mutate(STRIKE = ifelse(STRIKES>0,1,0),
         WEEK = factor(isoweek(DATE), levels=1:53),
         MONTH = factor(month(DATE), levels=1:12))

# d_day_hour$WEEK <- as.factor(d_day_hour$WEEK)
# d_day_hour$MONTH<- as.factor(d_day_hour$MONTH)

### Set 3: Data aggregated by day
### - every line correspondes to a particular date with STRIKES = num strikes, and STRIKE = 1/0 
d_day <- d_day_hour %>%
  group_by(`AIRPORT ID`,DATE) %>%
  summarise(`FLTS TOTAL` = sum(`FLTS TOTAL`), STRIKES = sum(STRIKE),
            BIRDCOUNT = first(BIRDCOUNT),
            TEMP = mean(TEMP,na.rm=TRUE), SLP = mean(SLP, na.rm=TRUE),
            `WIND ANGLE` = mean(`WIND ANGLE`, na.rm=T), `WIND SPEED`= mean(`WIND SPEED`,na.rm=TRUE),
            CLOUDH = mean(CLOUDH, na.rm=T), COVERH = mean(COVERH, na.rm=T),
            VIS = mean(VIS, na.rm=T),
            PRECIP = mean(PRECIP, na.rm=T)) %>%
  mutate(STRIKE = ifelse(STRIKES>0,1,0),
         WEEK = factor(isoweek(DATE), levels=1:53),
         MONTH = factor(month(DATE), levels=1:12))

### Quick - look
### Heavily unbalanced base set
table(d$STRIKE)

### High zero counts
# table(d_day_hour$STRIKES)
# table(d_day$STRIKES)
```

```{r}
# simple RF
d <- dn %>% filter(year(DATE) %in% 2010:2018) #%>% na.omit()

d <- as.data.frame(d)
d$STRIKE <- as.factor(ifelse(d$STRIKE==0, "NO", "YES"))


# Remove AirportId and Date columns
d <- subset(d, select = -c(`AIRPORT ID`, DATE)) %>% rename(WS = `WIND SPEED`, WA = `WIND ANGLE`, FLTS = `FLTS TOTAL`)

trainyn <- createDataPartition(d$STRIKE, p=0.8, list=FALSE)
dfTrain <- d[trainyn,]
dfTest <- d[-trainyn,]


rf1 <- randomForest::randomForest(STRIKE ~ BIRDCOUNT + FLIGHTCOUNT + TEMP + DEWP + SLP +
                                    VISIB + WDSP + MXSPD + PRCP + FOG + RAIN_DRIZZLE + SNOW_ICE + HAIL +
                                    THUNDER + MONTH + DAYOFWEEK + SEASON,
  data = dfTrain,
  mtry = 2,
  importance = TRUE,
  proximity = TRUE,
  do.trace = 100
)

rf1

dfTest$PREDICT <- predict(rf1, dfTest, type="class")
confusionMatrix(dfTest$PREDICT, dfTest$STRIKE, positive="YES")
# better on training set despite oob, out of sample/validation also 3.3% FN rate, 14/420
# balance issue?
nrow(dfTest[dfTest$STRIKE == "NO",])

# as.factor(STRIKE) ~ 
#                WS + WA + BIRDCOUNT + FLTS+ 
#                TEMP + COVERH + CLOUDH + VIS + SLP + PRECIP + MONTH

rf2 <- train(STRIKE ~ BIRDCOUNT + FLIGHTCOUNT + TEMP + DEWP + SLP +
                                    VISIB + WDSP + MXSPD + PRCP + FOG + RAIN_DRIZZLE + SNOW_ICE + HAIL +
                                    THUNDER + MONTH + DAYOFWEEK + SEASON,
            data=dfTrain, method="gbm",
            trControl = trainControl(method="cv", number=5, sampling="down"),
            tuneGrid = expand.grid(interaction.depth = 1:3,
                                   n.trees = (1:5)*50,
                                   shrinkage=0.1,
                                   n.minobsinnode = c(1,10,20)),
            verbose=FALSE)
rf2
varImp(rf2)

dfTest$PREDICT2 <- predict(rf2,dfTest)
confusionMatrix(dfTest$PREDICT2, dfTest$STRIKE, positive="YES")

### not great
rf3 <- train(STRIKE ~ BIRDCOUNT + FLIGHTCOUNT + TEMP + DEWP + SLP +
                                    VISIB + WDSP + MXSPD + PRCP + FOG + RAIN_DRIZZLE + SNOW_ICE + HAIL +
                                    THUNDER + MONTH + DAYOFWEEK + SEASON,
            data=dfTrain, method="AdaBoost.M1",
            trControl = trainControl(method="cv", n=5, sampling="down"))

dfTest$PREDICT3 <- predict(rf3,dfTest)
confusionMatrix(dfTest$PREDICT3, dfTest$STRIKE, positive="YES")


glmP <- glm(STRIKECOUNT ~ BIRDCOUNT + FLIGHTCOUNT + TEMP + DEWP + SLP +
                                    VISIB + WDSP + MXSPD + PRCP + FOG + RAIN_DRIZZLE + SNOW_ICE + HAIL +
                                    THUNDER + MONTH + DAYOFWEEK,
            data=dfTest, family=poisson)
summary(glmP)

g <- gam(STRIKECOUNT ~ BIRDCOUNT + FLIGHTCOUNT + TEMP + DEWP + SLP +
                                    VISIB + WDSP + MXSPD + PRCP + FOG + RAIN_DRIZZLE + SNOW_ICE + HAIL +
                                    THUNDER + MONTH + DAYOFWEEK, 
         method="REML", data=dfTrain)

summary(g)

par(mfrow = c(2,2))
gam.check(g)

g <- gam(STRIKECOUNT ~ s(as.integer(week(DATE)), bs = 'cc', k = 53) + s(as.integer(DATE)) +BIRDCOUNT + FLIGHTCOUNT + TEMP + DEWP + SLP +
                                    VISIB + WDSP + MXSPD + PRCP + FOG + RAIN_DRIZZLE + SNOW_ICE + HAIL +
                                    THUNDER, method="REML", data=dfTrain)
summary(g)
plot(g)
```


```{r}
d_day2 <- d_day %>% mutate(RATIO = STRIKES/`FLTS TOTAL`) %>% dplyr::filter(year(DATE) %in% 2014:2018)
d_day3 <- d_day %>% mutate(RATIO = STRIKES/`FLTS TOTAL`) %>% dplyr::filter(year(DATE) %in% 2014:2018, STRIKE==1)

ggplot(d_day2, aes(WEEK,RATIO)) + geom_boxplot() + geom_abline(slope=0,intercept=mean(d_day$RATIO))
m1 <- mean(d_day2$RATIO, na.rm=T)
s1 <- sd(d_day2$RATIO, na.rm=T)
m2 <- mean(d_day3$RATIO, na.rm=T)
s2 <- sd(d_day3$RATIO, na.rm=T)


# 0.0002658522
# 0.0004855315

lower <- m + s
upper <- m + 2*s

ggplot(d_day3, aes(RATIO)) + geom_histogram(bins=50) + geom_vline(xintercept=m1, colour="blue") + geom_vline(xintercept=m1+s1, colour="red") 
library(fitdistrplus)


d_day2 <- data %>% mutate(RATIO = STRIKECOUNT/FLIGHTCOUNT) %>% dplyr::filter(year(DATE) %in% 2000:2018)
x<- d_day2$RATIO
is
plotdist(d_day2$RATIO, demp=TRUE)
data %>% group_by(STRIKE) %>% summarise(n())
descdist(x[x>0], boot=1000)
summary(d_day2$RATIO)
y <- fitdist(x[x>0]/10, "gamma")
summary(y)
denscomp(y)
cdfcomp(y)
```



### Logist regression - DAY

```{r Logistic - Month dummies}
df <- d_day
table(df$STRIKES)
train_indicator <- createDataPartition(df$STRIKES, p = 0.8, list = FALSE)
dfTrain <- df[train_indicator,]
dfTest <- df[-train_indicator,]

#SLP CLOUDH
glmL <- glm(STRIKE ~ 
              `FLTS TOTAL` + BIRDCOUNT + TEMP + `WIND ANGLE` + 
              `WIND SPEED` + COVERH + VIS + PRECIP + MONTH,
                data=dfTrain, family=binomial, na.action=na.exclude)
summary(glmL)

classify <- function(threshold, data){ ifelse(data > threshold, 1 ,0) }

cglmL <- train(STRIKE ~ 
              `FLTS TOTAL` + BIRDCOUNT + TEMP + `WIND ANGLE` + 
              `WIND SPEED` + COVERH + VIS + PRECIP + MONTH,
              data=dfTrain, method="glm", family="binomial", na.action=na.exclude)
summary(cglmL)
# predict on test data
dfTest$PREDICT <- predict(glmL,dfTest, type="response")
dfTest$CLASS <- as.factor(classify(0.4,dfTest$PREDICT))

confusionMatrix(dfTest$CLASS,as.factor(dfTest$STRIKE), positive="1")
```


```{r Logistic / Mixed Model with Month as random effect (intercept)}

# Scaling - N(0,1), needed for mixed effects estimation as otherwise seems to have numeric issues
scales <- build_scales(dfTrain, cols = c("FLTS TOTAL","BIRDCOUNT","TEMP","SLP","WIND ANGLE", "WIND SPEED", "CLOUDH","COVERH","VIS","PRECIP"))
dfTrainS <- fastScale(dfTrain,scales)
glmLME <- glmer(STRIKE ~ 
              `FLTS TOTAL` + BIRDCOUNT + TEMP + `WIND ANGLE` + 
              `WIND SPEED` + COVERH + VIS + PRECIP + (1|MONTH),
                data=dfTrainS, family=binomial)
summary(glmLME)

phtest_glmer <- function (glmerMod, glmMod, ...)  {  ## changed function call
    coef.wi <- coef(glmMod)
    coef.re <- fixef(glmerMod)  ## changed coef() to fixef() for glmer
    vcov.wi <- vcov(glmMod)
    vcov.re <- vcov(glmerMod)
    names.wi <- names(coef.wi)
    names.re <- names(coef.re)
    coef.h <- names.re[names.re %in% names.wi]
    dbeta <- coef.wi[coef.h] - coef.re[coef.h]
    df <- length(dbeta)
    dvcov <- vcov.re[coef.h, coef.h] - vcov.wi[coef.h, coef.h]
    stat <- abs(t(dbeta) %*% as.matrix(solve(dvcov)) %*% dbeta)  ## added as.matrix()
    pval <- pchisq(stat, df = df, lower.tail = FALSE)
    names(stat) <- "chisq"
    parameter <- df
    names(parameter) <- "df"
    alternative <- "one model is inconsistent"
    res <- list(statistic = stat, p.value = pval, parameter = parameter, 
        method = "Hausman Test",  alternative = alternative,
                data.name=deparse(getCall(glmerMod)$data))  ## changed
    class(res) <- "htest"
    return(res)
}
phtest_glmer(glmLME, glmL)
```



```{r Poisson/NB}


glmL <- glm(STRIKE ~ 
              `FLTS TOTAL` + BIRDCOUNT + TEMP + SLP + `WIND ANGLE` + 
              `WIND SPEED` + CLOUDH + COVERH + VIS + PRECIP + MONTH + 
              VIS*`WIND SPEED`,
                data=dfTrain, family=binomial, na.action=na.exclude)
summary(glmL)

dfTrain$P <- predict(glmL,data=dfTrain,type="response",na.action = na.exclude)
dfTrain$LOGIT <- log(dfTrain$P/(1-dfTrain$P))

dx <- pivot_longer(dfTrain,c(`FLTS TOTAL`, BIRDCOUNT, TEMP , SLP , `WIND ANGLE` , 
              `WIND SPEED` , CLOUDH , COVERH , VIS , PRECIP),names_to="predictors", values_to="predictor.value")

ggplot(dx, aes(LOGIT,(predictor.value))) +
  geom_point(size=0.5, alpha=0.2)+
  geom_smooth(method="loess")+
  theme_bw()+facet_wrap(~predictors, scales="free_y")

ggplot(dfTrain, aes(COVERH,STRIKE)) + geom_point() + facet_wrap(~MONTH)

scales <- build_scales(dfTrain, cols = c("FLTS TOTAL","BIRDCOUNT","TEMP","SLP","WIND ANGLE", "WIND SPEED", "CLOUDH","COVERH","VIS","PRECIP"))
dfTrainS <- fastScale(dfTrain,scales)
glmLME <- glmer(STRIKE ~ 
              `FLTS TOTAL` + BIRDCOUNT + TEMP + SLP + `WIND ANGLE` + 
              `WIND SPEED` + COVERH + VIS + PRECIP + (1|MONTH) + 
              VIS*`WIND SPEED`,
                data=dfTrainS, family=binomial)
summary(glmLME)

dfTest <- dfTest %>%
  mutate(P = predict(glmL,dfTest,type="response"),
         PRED = ifelse(P >= log_threshold,1,0),
         ACC = ifelse(STRIKE==PRED,TRUE,FALSE)) 

p <- predict(glmL, dfTest, type="response")
pr <- prediction(p, dfTest$STRIKE)
prf <- performance(pr,measure="tpr", x.measure="fpr")
plot(prf)
performance(pr, measure="auc")
```

```{r}
glmNB <- glm.nb(STRIKES ~ 
                  `FLTS TOTAL` + BIRDCOUNT + TEMP + `WIND ANGLE` + 
                  `WIND SPEED` + CLOUDH + COVERH + VIS + PRECIP + MONTH,
                data=dfTrain)
summary(glmNB)

glmP <- glm(STRIKES ~ 
                  `FLTS TOTAL` + BIRDCOUNT + TEMP + `WIND ANGLE` + 
                  `WIND SPEED` + CLOUDH + COVERH + VIS + PRECIP + MONTH,
            data=dfTrain, family=poisson)
summary(glmP)

dfTest$PREDICTN <- predict(glmP,dfTest, type="response")

```


```
```{r}
# daily probability
df$PACT <- df$STRIKES/df$`FLTS TOTAL`

# daily probability across months
ggplot(df, aes(MONTH, PACT)) + geom_boxplot()

# daily P modelled directly
glmLp <- glm(PACT ~ 
              `FLTS TOTAL` + BIRDCOUNT + TEMP + SLP + `WIND ANGLE` + 
              `WIND SPEED` + CLOUDH + COVERH + VIS + PRECIP + MONTH + 
              VIS*`WIND SPEED`,
             data=df, 
             weights = `FLTS TOTAL`, family=binomial, na.action=na.exclude)
summary(glmLp)

df$PPRED <- predict(glmLp,df,type="response")
df$PPRED <- round(df$PPRED,5)
df$DELTA <- df$PACT-df$PPRED

ggplot(df, aes(log(COVERH),PACT)) + geom_point() + facet_wrap(~MONTH)

hm <- hurdle(STRIKES ~ 
              `FLTS TOTAL` + BIRDCOUNT + TEMP + SLP + `WIND ANGLE` + 
              `WIND SPEED` + CLOUDH + COVERH + VIS + PRECIP +
              VIS*`WIND SPEED`, data = dfTrainS, dist = "negbin", zero.dist="binomial")
summary(hm)
```

```{r VISUAL NALYSIS}
# Flight numbers over time
flights <- ggplot(df, aes(MONTH, `FLTS TOTAL`)) + geom_boxplot()

# Weather
temp <- ggplot(df, aes(MONTH, TEMP/10)) + geom_boxplot()
vis <- ggplot(df, aes(MONTH, VIS)) + geom_boxplot()
coverh <- ggplot(df,aes(MONTH, COVERH)) + geom_boxplot()

grid.arrange(flights,temp,vis, coverh, ncol=2)

dff <- d_day_hour

ggplot(dff, aes(VIS,STRIKES)) + geom_point() + facet_wrap(~MONTH)

#average characteristics of strike days
x <- dff %>% group_by(STRIKE, MONTH) %>% summarise(TEMP = mean(TEMP, na.rm = T),
                                       VIS = mean(VIS, na.rm = T),
                                       COVERH = mean(COVERH, na.rm=T))
ggplot(dff, aes(STRIKE,COVERH, group=STRIKE)) + geom_boxplot() + facet_wrap(~MONTH) 
```
```{r SVM}
make.grid = function(x, n = 75) {
  grange = apply(x, 2, range)
  x1 = seq(from = grange[1,1], to = grange[2,1], length = n)
  x2 = seq(from = grange[1,2], to = grange[2,2], length = n)
  expand.grid(X1 = x1, X2 = x2)
}

dsvm <- dplyr::select(dfTrainS, STRIKE, VIS,CLOUDH) %>% ungroup()


ggplot(dsvm, aes(x=(VIS),y=(CLOUDH),colour=STRIKE)) + geom_point()
svmm <- svm(STRIKE ~ ., kernel="polynomial", cost=10, data = dsvm, na.action=na.exclude, scale=FALSE)
plot(svmm,dsvm)
plopred <- predict(svmm,dfTrain)
plot(dfTrain$STRIKE,pred)
summary(svmm)
plot.svm(svmm)
```



```{r}

#ds <- dplyr::filter(d_day_hour,year(DATE) %in% 2014:2018)

# balance the data set
table(d_day_hour$STRIKE)
# SMOTE oversampling for Classification 
upsample.data <-
  smote(
    STRIKE ~ .,
    d_day_hour,
    perc.over = 5600,
    k = 10
  )

upsample.data <- upsample.data %>%
  mutate(STRIKE = ifelse(STRIKE == 0, "NO", "YES"))


# split the data
trainyn <- createDataPartition(d_day_hour$STRIKE,p=0.8,list=FALSE)
dtrain <- d_day_hour[trainyn,]
dtest <- d_day_hour[-trainyn,]

# control = trainControl(
#   method = 'cv',
#   number = 10,
#   classProbs = TRUE,
#   summaryFunction = twoClassSummary
# )
# 
# xgb <- train(STRIKE ~ `FLTS TOTAL` + BIRDCOUNT + TEMP + `WIND ANGLE` + 
#             `WIND SPEED` + SLP + CLOUDH + COVERH + VIS + MONTH,
#              data = dtrain,
#              method = "xgbTree",
#              metric = "Sens",
#              trControl = control
# )

tm <- rpart(STRIKE ~ `FLTS TOTAL` + BIRDCOUNT + TEMP + `WIND ANGLE` + 
            `WIND SPEED` + SLP + CLOUDH + COVERH + VIS + MONTH, data=dtrain)
pred = predict(tm, type="class")

gm <- gbm(STRIKE ~ `FLTS TOTAL` + BIRDCOUNT + TEMP + `WIND ANGLE` + 
            `WIND SPEED` + SLP + CLOUDH + COVERH + VIS + MONTH, 
          data=dtrain, n.trees = 1000)

confusionMatrix(predict(gm, dtest, n.trees=1000), dtest$STRIKE)
```


```{r Logistic regression - DAY/HOUR}

m_day_hour_notime <- glm(STRIKE ~ `FLTS TOTAL` + `WIND SPEED` + `WIND ANGLE` + TEMP + SLP + BIRDCOUNT + 
                       CLOUDH + COVERH + VIS, data = d_day_hour,family=binomial(link='logit'))
summary(m_day_hour_notime)

m_day_hour_wfactor <- glm(STRIKE ~ `FLTS TOTAL` + `WIND SPEED` + `WIND ANGLE` + TEMP + SLP + BIRDCOUNT + 
                       CLOUDH + COVERH + VIS + WEEK, data = d_day_hour,family=binomial(link='logit'))
summary(m_day_hour_wfactor)

m_day_hour_mfactor <- glm(STRIKE ~ `FLTS TOTAL` + `WIND SPEED` + `WIND ANGLE` + TEMP + SLP + BIRDCOUNT + 
                       CLOUDH + COVERH + VIS + MONTH, data = d_day_hour,family=binomial(link='logit'))
summary(m_day_hour_mfactor)

### too much noise in hourly information?
```

```{r Logistic regression - DAY}

m_day_notime <- glm(STRIKE ~ `FLTS TOTAL` + `WIND SPEED` + `WIND ANGLE` + TEMP + SLP + BIRDCOUNT + 
                       CLOUDH + COVERH + VIS, data = d_day,family=binomial(link='logit'))
summary(m_day_notime)

m_day_wfactor <- glm(STRIKE ~ `FLTS TOTAL` + `WIND SPEED` + `WIND ANGLE` + TEMP + SLP + BIRDCOUNT + 
                       CLOUDH + COVERH + VIS + WEEK, data = d_day,family=binomial(link='logit'))
summary(m_day_wfactor)

m_day_mfactor <- glm(STRIKE ~ `FLTS TOTAL` + `WIND SPEED` + `WIND ANGLE` + TEMP + SLP + BIRDCOUNT + 
                       CLOUDH + COVERH + VIS + MONTH, data = d_day,family=binomial(link='logit'))
summary(m_day_mfactor)

```

```{r Count Models - DAY}

m <-glmer.nb(STRIKES ~ `FLTS TOTAL` +  (1|MONTH), data = d_day)

m <-glm.nb(STRIKES ~`FLTS TOTAL` + VIS + TEMP + COVERH + CLOUDH + SLP + BIRDCOUNT +
             `WIND SPEED` + `WIND ANGLE` + as.factor(WEEK), data = d_day_hour)

m_count_day_notime <- hurdle(STRIKES ~ SLP + BIRDCOUNT + `FLTS TOTAL` + `WIND SPEED` + `WIND ANGLE` + TEMP + 
                              CLOUDH + COVERH + VIS, 
                               data = d_day, dist="negbin")
m_count_day_notime <- glm.nb(STRIKES ~ `FLTS TOTAL` + `WIND SPEED` + `WIND ANGLE` + TEMP + 
                                 SLP + BIRDCOUNT + CLOUDH + COVERH + VIS, 
                               data = d_day)
summary(m_count_day_notime)
#### Seems that models very sensitive to scale of variables ...
#### When using hurdle/zeroinfl with "negbin", it seems that theta goes very large, which would indicate that poisson model can be used (as it woudl approximate it) -> same above
# m_count_day_mfactor <- hurdle(STRIKES ~ `FLTS TOTAL` + `WIND SPEED` + `WIND ANGLE` + TEMP + 
#                                  SLP + BIRDCOUNT + CLOUDH + COVERH + VIS + MONTH,
#                                data = dddd, dist="poisson")

# m_count_day_mfactor <- glmer.nb(STRIKES ~ `FLTS TOTAL` + `WIND SPEED` + `WIND ANGLE` + TEMP + 
#                                  SLP + BIRDCOUNT + CLOUDH + COVERH + VIS + (1|MONTH),
#                                data = dddd)

### Even after scaling .. still large theta
ds_day <- fastScale(d_day)
ds_day$STRIKES <- d_day$STRIKES
m_count_day_wfactor <- glm(STRIKES ~ `FLTS TOTAL` + `WIND SPEED` + `WIND ANGLE` + TEMP + 
                                 SLP + BIRDCOUNT + CLOUDH + COVERH + VIS + WEEK, family= "poisson",
                               data = ds_day)
m_count_day_wfactor <- glm.nb(STRIKES ~ `FLTS TOTAL` + `WIND SPEED` + `WIND ANGLE` + TEMP + 
                                 SLP + BIRDCOUNT + CLOUDH + COVERH + VIS + WEEK,
                               data = ds_day)
summary(m_count_day_wfactor)
        
### Theta smaller, so Nb probably better       
m_count_day_mfactor <- glm(STRIKES ~ `FLTS TOTAL` + `WIND SPEED` + `WIND ANGLE` + TEMP + 
                                 SLP + BIRDCOUNT + CLOUDH + COVERH + VIS + MONTH, 
                               data = ds_day, family="poisson")
m_count_day_mfactor <- glm.nb(STRIKES ~ `FLTS TOTAL` + `WIND SPEED` + `WIND ANGLE` + TEMP + 
                                 SLP + BIRDCOUNT + CLOUDH + COVERH + VIS + MONTH, 
                               data = ds_day)
m_count_day_mfactor <- glmer.nb(STRIKES ~ `FLTS TOTAL` + `WIND SPEED` + `WIND ANGLE` + TEMP + 
                                 SLP + BIRDCOUNT + CLOUDH + COVERH + VIS + 
                                  (1+ VIS|MONTH) , 
                               data = ds_day)


summary(m_count_day_mfactor)



```


```{r Trees?}

tm <- gbm(STRIKES ~`FLTS TOTAL` + `WIND SPEED` + `WIND ANGLE` + TEMP + 
                                 SLP + BIRDCOUNT + CLOUDH + VIS, data = d_day)

```


dx <- d %>% dplyr::filter(DATE == "2014-01-01") 

%>% group_by(`AIRPORT ID`, DATE) %>% 
  summarise(RTIME=mean(RTIME),STRIKES = sum(STRIKE, na.rm=T), `FLTS TOTAL`=sum(`FLTS TOTAL`, na.rm=T), 
            TEMP = mean(TEMP,na.rm=TRUE),
            `WIND ANGLE` = mean(`WIND ANGLE`, na.rm=T), `WIND SPEED`= mean(`WIND SPEED`,na.rm=TRUE),
            BIRDS = sum(BIRDCOUNT), CLOUDH = mean(CLOUDH, na.rm=T))

day_strike_binary <- d %>% dplyr::filter(year(DATE) %in% 2014:2019) %>%
  group_by(`AIRPORT ID`, DATE) %>% 
  summarise(RTIME=mean(RTIME),STRIKES = sum(STRIKE, na.rm=T), `FLTS TOTAL`=mean(`FLTS TOTAL`, na.rm=T), 
            TEMP = mean(TEMP,na.rm=TRUE),
            `WIND ANGLE` = mean(`WIND ANGLE`, na.rm=T), `WIND SPEED`= mean(`WIND SPEED`,na.rm=TRUE),
            BIRDS = sum(BIRDCOUNT), CLOUDH = mean(CLOUDH, na.rm=T))
  
day_strike_binary <- day_strike_binary %>% mutate(YN = if_else(STRIKES>0,1,0))

m <- glm(YN ~ `FLTS TOTAL` + `WIND SPEED` + TEMP + `WIND ANGLE` + BIRDS + CLOUDH + as.factor(week(DATE)),
         data = day_strike_binary,family=binomial(link='logit'))
mzi <- zeroinfl(STRIKES ~ `FLTS TOTAL` + `WIND SPEED` + TEMP + `WIND ANGLE` + BIRDS + CLOUDH + as.factor(week(DATE)),
         data = day_strike_binary)

### DON'T RUN ANY OF THIS ...  takes too long and has issues

m <- glm.nb(STRIKE ~ `FLTS TOTAL` + `WIND SPEED` + TEMP, data = d, control = list(maxit = 25, epsilon=0.000001, trace=FALSE))


d_byday <- d %>% dplyr::filter(year(DATE) %in% 2014:2019) %>%
  group_by(`AIRPORT ID`, DATE) %>% 
  summarise(RTIME=mean(RTIME),STRIKES = sum(STRIKE, na.rm=T), `FLTS TOTAL`=sum(`FLTS TOTAL`, na.rm=T), 
            TEMP = mean(TEMP,na.rm=TRUE),`WIND SPEED`=mean(`WIND SPEED`,na.rm=TRUE))

m <- glm.nb(STRIKES ~ `FLTS TOTAL` + TEMP + as.factor(week(DATE)), data = dbyday)
mh <- hurdle(STRIKE ~ `FLTS TOTAL` + TEMP + as.factor(week(DATE)), data = dplyr::filter(d,year(DATE) %in% 2000:2010))

model <- glm(Survived ~.,family=binomial(link='logit'),data=train)
```


```{r Classifier - Forest GBM}

```



# Count Models

```{r Initial Analysis}

# Number of strikes per week (note this is different from the ave strikes per week aggregated by days)

master <- readRDS("KDEN complete.RDS")
mw <- master %>% 
  dplyr::filter(DATE > "1994-08-01") %>% 
  mutate(WEEK = isoweek(DATE)) %>% 
  group_by(YEAR=year(DATE), WEEK) %>% 
  summarise(x=sum(STRIKE)) %>% 
  group_by(WEEK) %>% 
  summarise(NWEEKS=n(),STRIKES=sum(x), RATIO=STRIKES/NWEEKS)


# Map back to master file
master <- master %>% mutate(WEEK = isoweek(DATE))
mw <- dplyr::select(mw, WEEK, RATIO)
master <- left_join(master,mw, by=c("WEEK"="WEEK"))

x <- master %>% group_by(WEEK) %>% summarise(RATIO = mean(RATIO))
mean(x$RATIO)
sd(x$RATIO)


master <- master %>% group_by(DATE, RTIME) %>% summarise
m <- lm(RATIO ~ TEMP + as.factor(week(DATE)), data=master)

# 53 weeks? Seems part of ISO standard because Jan 1st not always a Monday
mw <- master %>% group_by(DATE, WEEK = week(DATE)) 

## Calculate the probability of a strike ... WE SEEM TO LOSE SOME ROWS IN THIS AGGREGATION - SAME RUNWAY STRIKES?
# NOTE, CAN'T USE n() as it counts records and thus includes days with no strikes (=NA)
masterl <- master %>% 
  group_by(`AIRPORT ID`, DATE,TD, RUNWAY) %>%
  summarise(STRIKES = sum(STRIKE), TOTAL = sum(TOTAL), ARRIVALS = sum(ARRIVALS), DEPARTURES = sum(DEPARTURES),
            STRIKEP = STRIKES/TOTAL)
```

```{r Graphs}
x <- master %>% group_by(WEEK) %>% summarise(RATIO = mean(RATIO))
ggplot(x, aes(x=WEEK,y=RATIO, group=WEEK)) + geom_boxplot()

master <- master %>% mutate(WEEK = week(DATE))

m <- gbm(STRIKE ~ WEEK + TOTAL,data=master)

```

```{r}
ggplot(d_day, aes(TEMP,VIS, colour=STRIKE)) + geom_point()

d_day <- transform(d_day, group=cut(d_day$TEMP,c(-1000,0,100,200,300,400), include.lowest=TRUE))

d_day %>% group_by(group) %>% summarise(N=n(), S = sum(STRIKES), R=S/N)
                             
glmL <- glm(STRIKE ~ group +
              `FLTS.TOTAL` + BIRDCOUNT + TEMP + `WIND.ANGLE` + 
              `WIND.SPEED` + COVERH + VIS + PRECIP + MONTH,
                data=d_day, family=binomial, na.action=na.exclude)
summary(glmL)

classify <- function(threshold, data){ ifelse(data > threshold, 1 ,0) }



ggplot(d_day, aes(MONTH, `WIND ANGLE`)) + geom_boxplot()


```

```{r}
library(rpart)
library(rpart.plot)

x <- dfTrain %>% ungroup() %>% 
  select(-c(DATE, `AIRPORT ID`)) %>% na.omit() %>% 
  mutate(TARGET = as.factor(STRIKE))

# flaky results
#dfTrainBalanced <- smote(STRIKE ~ ., data=x, perc.over=2)

# dfTrainBalancedUp <- upSample(x %>% select(-CLASS),x$CLASS)
# table(dfTrainBalancedUp$Class)
# table(x$Class)

t <- rpart(STRIKE ~
              `FLTS TOTAL`+`WIND ANGLE`+ SLP + BIRDCOUNT +
              `WIND SPEED` + COVERH + VIS + PRECIP, data=dfTrainBalancedUp, control=rpart.control(minsplit=2))

prp(t)

rf <- train(TARGET ~
              `FLTS TOTAL`+`WIND ANGLE`+ SLP + TEMP + BIRDCOUNT +
              `WIND SPEED` + COVERH + CLOUDH + VIS + PRECIP + MONTH,
            data=x, method="gbm",
            trControl = trainControl(method="cv", number=5, sampling="smote"),
            tuneGrid = expand.grid(interaction.depth = 1:10,
                                   n.trees = (1:5)*50,
                                   shrinkage=0.1,
                                   n.minobsinnode = c(1,10,20)),
            verbose=FALSE)
rf
varImp(rf)

predict(rf,dfTest)

ada <- train(TARGET ~`FLTS TOTAL`+`WIND ANGLE`+ SLP + TEMP + BIRDCOUNT +
              `WIND SPEED` + COVERH + CLOUDH + VIS + PRECIP + as.numeric(MONTH),
            data=x, method="adaboost",
            trControl = trainControl(method="cv", number=5, savePredictions = "all"))

x <- d_day
x$RATIO <- x$STRIKES/x$`FLTS TOTAL`
ggplot(x, aes(y=RATIO, group=month(DATE))) + geom_boxplot()

x <- x %>% ungroup() %>% dplyr::select(-c(`AIRPORT ID`)) 
x <- x %>% mutate(MONTH = as.numeric(MONTH),WEEK = as.numeric(WEEK), STRIKE = as.numeric(STRIKE))
x <- x[,-ncol(x)]
x <- data.frame(scale(x))
pc <- princomp(as.matrix(na.omit(x)), cor=T)
plot(pc, type="l")
summary(pc)
loadings(pc)

pcv <- prcomp(na.omit(x))
pcv1 <- data.frame(pcv$x[,1:5])
plot(pcv1, pch=16, col=rgb(0,0,0,0.5))
k <- kmeans(pcv1,2,nstart=2)
plot(pcv1, col=k$clust, pch=16)
```

